\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage[paperwidth=8.5in,paperheight=11in]{geometry}
\geometry{verbose,tmargin=0.5in,bmargin=1in,lmargin=1in,rmargin=1in}
\providecommand{\alert}[1]{\textbf{#1}}

\title{\textbf{Youngstown State University}}
\author{Department of Mathematics and Statistics}
\date{Grading Policy for STAT 5840, Summer 2011}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.02}}

\begin{document}

\maketitle


\section*{Contact Information:}
\label{sec-1}


\begin{center}
\begin{tabular}{l|l}
\hline
 \textbf{Instructor:}    &  G. Jay Kerns, Ph.D.                                                          \\
 \textbf{Office:}        &  620 Lincoln Building                                                         \\
 \textbf{Phone No.:}     &  (330) 941-3310                                                               \\
 \textbf{E-mail:}        &  \href{mailto:gkerns@ysu.edu}{gkerns@ysu.edu}                                 \\
 \textbf{Webpage:}       &  \texttt{http://personal.ysu.edu/\textasciitilde{}gkerns}                     \\
 \textbf{Course page:}   &  \href{http://www.moodle.ysumathstat.org}{http://www.moodle.ysumathstat.org}  \\
 \textbf{Office hours:}  &  15:00 – 16:00 MWF, or by appointment.                                        \\
\hline
\end{tabular}
\end{center}
\section*{Graded Assessments:}
\label{sec-2}


\begin{center}
\begin{tabular}{l|l}
\hline
 Homework and In-class Assignments  &  300 points  \\
\hline
 \textbf{Total Points Possible}     &  300 points  \\
\hline
\end{tabular}
\end{center}
\section*{Final Grade Assignment:}
\label{sec-3}


\begin{center}
\begin{tabular}{lllll}
\hline
 A          &  B          &  C          &  D          &  F          \\
\hline
 270 - 300  &  240 - 269  &  210 - 239  &  180 - 209  &  below 180  \\
\hline
\end{tabular}
\end{center}
\section*{Make-up policy:}
\label{sec-4}

\textbf{Late homework or projects} will be penalized 10\% of the grade per day. 
\section*{Homework and projects:}
\label{sec-5}

Exercises will be assigned during class. Some of these will be collected. It may become necessary to do some of these exercises during class; if so, I will let you know ahead of time. Grades will be based on performance on these exercises. 
\section*{Attendance:}
\label{sec-6}

It is expected that you will make every possible effort to attend all classes. Although I feel that it is essential for you to attend class, no direct punishment will be assessed to your final grade if you choose not to attend class. Students who miss class are responsible for finding out any pertinent information concerning the course from the instructor. 
\section*{Recommendations:}
\label{sec-7}

\begin{itemize}
\item Successful students have found that the subject demands frequent intensive study with lots of problem solving.
\item Participation in group discussions is often helpful to yourself and others in learning.
\item If you have any special problems or learning difficulties, please contact me immediately.
\end{itemize}

\end{document}