% Created 2011-08-25 Thu 13:05
\documentclass[11pt,english]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{color}
\usepackage{listings}
\input{handoutformat}
\providecommand{\alert}[1]{\textbf{#1}}

\title{Assignment: Evaluating Tests via Monte Carlo Methods}
\author{G. Jay Kerns}
\date{}

\begin{document}

\maketitle


\section*{Setup}
\label{sec-1}


Suppose $X_{1}$, $X_{2}$, \ldots{}, $X_{n}$ are a random sample from a distribution $f$ with mean $\mu$.  The standard two-sided test of the null hypothesis $H_{0}: \mu = 0$ is based on the statistic
\[
T = \frac{\Xbar}{S/\sqrt{n}}.
\]
At level $\alpha = 0.05$, one rejects $H_{0}$ if $|T| > t_{\alpha/2}$, where $t_{\alpha/2}$ is the percentile of a \texttt{t(df = n-1)} distribution with upper tail probability 0.025.  For this exercise, assume $n = 10$.


\begin{description}
\item[KNOW:] An implicit assumption for the above test is that the underlying population is Normal. What happens if that assumption is violated?
\end{description}
\section*{Questions.}
\label{sec-2}


\begin{itemize}
\item Using Monte Carlo simulation, approximate the level of this test when $f$ is
\begin{enumerate}
\item N(0,1),
\item Chi-square with 4 \texttt{df},
\item Uniform(0,1), and
\item Standard Logistic, Logis(0,1).
\end{enumerate}
\item For each population distribution, give a) the estimate of the level, b) a standard error for your estimate, and c) a paragraph stating what you have found.
\item After the four parts are finished, write a report collecting and summarizing your findings.  What have you learned about the robustness of the \emph{t}-test to departures from the normality assumption?
\end{itemize}

\textbf{Hints:}

\begin{itemize}
\item Download \texttt{robust.R} and modify it to suit your needs by changing the simulated data and the \texttt{t.test} command.
\end{itemize}

\end{document}