% Created 2011-08-26 Fri 16:02
\documentclass[11pt,english]{article}

\usepackage{color}
\usepackage{listings}
\input{handoutformat}
\providecommand{\alert}[1]{\textbf{#1}}

\title{Assignment: Importance Sampling and Antithetic Variables}
\author{STAT 5840 Summer 2011}
\date{}

\begin{document}

\maketitle


\section*{Questions}
\label{sec-1}



\begin{enumerate}
\item Obtain a Monte Carlo estimate $\hat{\theta}$ of 
   \[
   \int_{1}^{\infty} \frac{x^{3}}{\sqrt{2\pi}}\me^{-x^2/2} \diff x
   \]
   by importance sampling,  and also estimate the variance of $\hat{\theta}$.  \textbf{Note:} there are \emph{many} ways to do this problem, and I'm only looking for one (1) way.  Surprise me.
\item Compute a Monte Carlo estimate $\hat{\theta}$ of 
   \[
   \theta = \int_{0}^{2} \me ^{-x}\,\diff x
   \]
   by sampling from a Unif(0,2) distribution, and also estimate the variance of $\hat{\theta}$. Find another Monte Carlo estimator $\hat{\theta}^{\ast}$ by sampling from an Exp(1) distribution.  Which of the variances (of $\hat{\theta}$ and $\hat{\theta}^{\ast}$) is smaller, and why?
\item Use Monte Carlo integration with antithetic variables to estimate
   \[
   \theta = \int_{0}^{1} \frac{\me ^{-x}}{1 + x^{2}}\,\diff x,
   \]
   and find the approximate reduction in variance as a percentage of the variance without variance reduction (this means you also need to estimate $\theta$ the naive way and compare variances).
\end{enumerate}

\textbf{Hints:}


\begin{enumerate}
\item Look for an importance function which is easy to simulate and has heavier tails than $f$ (but still has finite variance).  You can also manipulate the integral beforehand with Calculus tricks which may give you something easier/better.  For hints how to set up your program, look at \texttt{explognormal.R}.
\item Not really much to say here.  Simulate some variates, compute some values, and find the \texttt{mean}.  For the variance you are looking at $\hat{\theta}(1 - \hat{\theta})/n$.
\item Take a look at the range of integration.  Do you see the natural choice of variates to simulate?  Now, remember from class: what's the easiest way to introduce correlation with variates like those?
\end{enumerate}

\end{document}